Group Signatures Meet Anonymous Credentials: Or, How to Build Privacy Preserving Identities with Accountability

### 2022-02-20 prism-backend/research/gsac/uas-model.tex
Working on non-frameability. Realized there are issues in the current definition of traceability.

\todo{Are generalizations where different groups can share issuer or inspector
  straight forward?} \\
\todo{The notation for multisets is not strictly correct. Still, leaving it as
  is for now.} \\
\commentwho{Jesus}{I'm trying to define \Sign, \Verify, \Inspect and \Judge such
  that multiple credentials can be used to produce a single signature, proving
  claims over arbitrary subsets of all the credentials owned by a user. \Verify
  and, consequently, \Inspect and \Judge need to receive all the \gpk's.
  \Inspect only requires one inspector's secret key, though, and will thus only
  produce that inspector's
  function of the signature and message (this is to avoid complex interactive
  protocols between many inspectors). This, however, creates a quite annoying
  syntax. It may just be easier to restrict signatures to claims over the
  attributes contained in \emph{one} credential.}

@@

\todo{Domains (attributes, messages, etc.) are not correctly defined here. But
  for now, just leave it as is to at least convey the idea. Also, this needs
  refinement in general: what properties (if any) do these functions need to
  meet in order to be meaningful?}
\begin{description}
\item[$\fissue: (\uattrs,\dattrs) \rightarrow 0/1$.] Governs what customized
  conditions an issuer requires in order to issue credentials, when receiving
  a request from a user with a set of disclosed attributes \dattrs, and
  undisclosed attributes \uattrs (which need to be kept private from the
  issuer). \fissue returns 1 to accept a request, 0 to reject it.  
\item[$\feval: (\msg,\sig) \rightarrow 0/1$.] Governs customized conditions upon
  which a verifier should accept (1) or reject (0) a given message-signature
  pair.
\item[$\finsp: (\upk,\scred) \rightarrow \bin^*$.] Defines what utility value,
  derived from the user's public key and credentials, should be extractable by
  the group inspector.
\end{description}

@@

    \caption{Detailed oracles available in our model (1/2). Oracles for
      generating key material for users, issuers, and inspectors.\\
      \doubt{Do we need to specify the functions in the GEN Oracles? They do
        not seem necessary in \SIGN/\CHALb/\INSPECT. Should they? Maybe there
        is something wrong here... Maybe they are needed in the games...}}
    \label{fig:oracles1}
  \end{figure*}
}

@@

        \procedure{$\CHALb(\msg,
          (\uid_{0,1},\uattrs_{0,1},\dattrs_{0,1},\ldblbrace (\sgid,\cscid_{0,1},\tuattrs_{0,1},\tdattrs_{0,1}) \rdblbrace))$}{%
          \pcif \cuid_0 \notin \HU \lor \cuid_1 \notin \HU: \pcreturn \bot \\
          \pccomment{Note: since both \uid's need to be honest, we must have
            all the}\\
          \pccomment{undisclosed attributes in the \ATTR lists.} \\
          \dattrs' \gets \DATTR[\cscid_0] \cap \DATTR[\cscid_1] \\
          \uattrs' \gets \UATTR[\cscid_0] \cap \UATTR[\cscid_1] \\
          \pcif (\dattrs_0 \cap \dattrs_1) \nsubseteq \dattrs':
          \pcreturn \bot \\
          \pcif (\uattrs_0 \cap \uattrs_1) \nsubseteq \uattrs':          
          \pcreturn \bot \\
          \csig \gets \Sign(\PRVUK[\uid_b],\msg,\uattrs_b,\dattrs_b, \\
          \hspace*{42pt}
          \ldblbrace (\GK[\sgid],\CRED[\scid_b],\tuattrs_b,\tdattrs_b) \rdblbrace) \\
          \CSIG \gets \CSIG \cup
          \lbrace (\csig,\msg,\scid_b,\uattrs_b,\dattrs_b,\sgid) \rbrace \\
          \pcreturn \csig
        }

@@

\paragraph{Anonymity.} %
In group signatures, anonymity captures that no adversary must be able to learn,
from any group signature, anything about its signer. In anonymous credentials,
it requires that no adversary should learn anything about the holder of a
successfully shown credential, beyond that he owns a credential attesting for
the claimed function of attributes it contains. In both GS and AC, it is also
typically required that
multiple signatures/presentations by the users are unlinkable. The approach to
formally state this property is in both cases frequently the same: the adversary
picks two (honest) users (or credentials in the AC case), the game randomly
chooses one of them, and lets the adversary request challenge
signatures/presentations from it. The adversary wins if it succeeds in guessing
which was the chosen user/credential better than guessing at random. In group
signatures, the game must also restrict the adversary from opening challenge
signatures. In anonymous credentials, the game must further constraint the
adversary to output credentials that have some common function of (a subset of)
its attributes.

In our notion of anonymity for \UAS, we need to merge the previous constraints.
Furthermore, a key difference with group signatures is that the game requires
the adversary to output credential identifiers, rather than user identifiers.
Specifically, this means that the adversary may actually output two credentials
that belong to the same user. Therefore, in some sense, the anonymity we get is
more general than that of group signatures. Moreover, in order to prevent
trivial wins by the adversary, we have to restrict that the challenge signatures
belong to a group that has been ``programmed'' with the same evaluation and
inspection functions. The formal specification of the anonymity game is given in
\figref{fig:exp-uas-anonb}.

@@

\begin{definition}{(Anonymity of \GSAC)}
  We define the advantage \AdvAnon of $\adv$ against \ExpAnonb as
  $\AdvAnon=|\Pr\lbrack\ExpAnono(1^\secpar)=1\rbrack-
  \Pr\lbrack\ExpAnonz(1^\secpar)=1\rbrack|$.
  %
  A \GSAC scheme satisfies anonymity if, for any p.p.t. adversary $\adv$,
  \AdvAnon is a negligible function of $1^\secpar$.
\end{definition}

\paragraph{Discussion on the generic anonymity of \UAS schemes.} %
This notion of anonymity is also more general in the sense that calls to
the \INSPECT oracle reveal an arbitrary function of the identity of the user
-- which can certainly be the identity itself, as in group signatures, ore any
other function computable from the \textcolor{red}{user public key and all of
  its credentials}. Of course, this requires to restrict the adversary to
programming equivalent inspection functions in the challenge signatures it
gets (or, more precisely, into the groups they belong to). While this could
appear as having a more restricted type of anonymity -- and indeed it could be
for some subset of inspection functions -- we point out that it is just a
generalization, as the concrete inspection function depends on the needs of
the use case where a \UAS scheme is to be integrated.

\paragraph{Traceability.} %
Traceability is one of the unforgeability-related properties in group
signatures. It captures that any signature accepted by \Verify needs to open
to one of the users that joined the group. While there is no traceability notion
in anonymous credentials, it is natural to map it to their unforgeability
property; if only because both require the issuer to be honest. Unforgeability
in anonymous credentials typically ensures that no adversary can get a verifier
to accept a credential presentation requiring a set of attributes that is not
contained in one of the credentials controlled by the adversary.
%
Our notion of traceability for \UAS...

\begin{figure}[htp!]
  \procedure{$\ExpTrace(1^\secpar)$}{%
     \parm \gets \Setup(1^\secpar) \\
     (\ipk,\isk) \gets \IKeyGen(\parm);~(\opk,\osk) \gets \OKeyGen(\parm);~ 
     \gpk \gets (\ipk,\opk) \\
     (\sig,\dattrs,\msg) \gets
     \adv^{\HUGEN,\CUGEN,\OBTISS,\ISSUE,\SIGN,\OPEN}(\gpk,\osk) \\
     \pcreturn 1~\pcif \Verify(\gpk,\sig,\dattrs,\msg) = 1~\land~( \\
     \pcind \pccomment{All \Obtain transcripts fail to open \sig} \\
     \pcind \forall \utrans \in \trans:
     \Open(\gpk,\osk,\utrans,\sig,\dattrs,\msg) = \bot~\lor \\
     \pcind \pccomment{Such transcript exists, but ...} \\
     \pcind \exists~\utrans~\st~\Open(\gpk,\osk,\utrans,\sig,\dattrs,\msg) =
     (\upk,\cred,\oproof)~\land (\\
     \pcind \pcind \pccomment{... the opening proof is rejected by \Judge ...} \\
     \pcind \pcind \Judge(\gpk,\upk,\cred,\oproof,\sig,\dattrs,\msg) = 0~\lor \\
     \pcind \pcind \pccomment{... or \Judge accepts the proof, but \dattrs was
       not issued to any credential} \\
     \pcind \pcind \dattrs \nsubseteq \ATTR[\cred] \\
     \pcind ) \\
     ) \\
     \pcreturn 0
  }
  \caption{Traceability experiment for \GSAC schemes.}
  \label{fig:exp-gsac-trace}
\end{figure}

\begin{definition}{(Traceability of \GSAC)}
  We define the advantage \AdvTrace of $\adv$ against \ExpTrace as
  $\AdvTrace=\Pr\lbrack\ExpTrace(1^\secpar)=1\rbrack$.
  %
  A \GSAC scheme satisfies traceability if, for any p.p.t. adversary $\adv$,
  \AdvTrace is a negligible function of $1^\secpar$.
\end{definition}

@@

\begin{figure}[htp!]
  \procedure{$\ExpNonframe(1^\secpar)$}{%
    \parm \gets \Setup(1^\secpar) \\
     (\ipk,\isk) \gets \IKeyGen(\parm);~(\opk,\osk) \gets \OKeyGen(\parm);~ 
     \gpk \gets (\ipk,\opk) \\
     (\sig,\dattrs,\msg,\upk,\cred,\oproof) \gets
     \adv^{\HUGEN,\CUGEN,\OBTAIN,\SIGN}(\gpk,\isk,\osk) \\
     \pcreturn 1~\pcif \Verify(\gpk,\sig,\dattrs,\msg) = 1 \land
     \OWNR[\cred] \in \HU \\%\land \cred \notin \CCRED~\land\\
     \pcind \Judge(\gpk,\upk,\cred,\oproof,\sig,\dattrs,\msg) = 1~\land \\
     \pcind \pccomment{\sig was not produced by \SIGN, or \upk and \cred do not match} \\
     \pcind (\sig \notin \SIG[\OWNR[\cred]] \lor \PUBUK[\OWNR[\cred]] \neq \upk) \\
     \pcreturn 0
  }
  \caption{Non-frameability experiment for \GSAC schemes.}
  \label{fig:exp-gsac-frame}
\end{figure}

>>>>

\todo{Are generalizations where different groups can share issuer or inspector
  straight forward?}

\todo{The notation for multisets is not strictly correct. Still, leaving it as
  is for now.}

\commentwho{Jesus}{I'm trying to define \Sign, \Verify, \Inspect and \Judge such
  that multiple credentials can be used to produce a single signature, proving
  claims over arbitrary subsets of all the credentials owned by a user. \Verify
  and, consequently, \Inspect and \Judge need to receive all the \gpk's.
  \Inspect only requires one inspector's secret key, though, and will thus only
  produce that inspector's
  function of the signature and message (this is to avoid complex interactive
  protocols between many inspectors). This, however, creates a quite annoying
  syntax. It may just be easier to restrict signatures to claims over the
  attributes contained in \emph{one} credential.}

\commentwho{Jesus}{Maybe we could remove the attribute sets \uattrs, \dattrs,
  \tuattrs, \tdattrs, from the syntax, and assume that they are part of the
  signatures and/or credentials. This would certainly lead to a cleaner syntax,
  but, what impact would that have in other definitions, etc? Specifically, can
  we support selective disclosure credentials without them? (if so, \fissue,
  \feval and \finsp would be essential in doing so.); Right now, I am more in
  favour of tryint to remove the attribute sets from the definitions, as much
  as possible (and see if we can still support plain selective disclosure).}

@@

\todo{Domains (attributes, messages, etc.) are not correctly defined here. But
  for now, just leave it as is to at least convey the idea. Also, this needs
  refinement in general: what properties (if any) do these functions need to
  meet in order to be meaningful?}
\begin{description}
\item[$\fissue: (\uattrs,\dattrs) \rightarrow 0/1$.] Governs what customized
  conditions an issuer requires in order to issue credentials, when receiving
  a request from a user with a set of disclosed attributes \dattrs, and
  undisclosed attributes \uattrs (which need to be kept private from the
  issuer). \fissue returns 1 to accept a request, 0 to reject it.  
\item[$\feval: (\msg,\scred) \rightarrow 0/1$.] Governs customized conditions
  to define whether the owner of a set of credentials -- or, more specifically,
  credentials encoding certain attributes -- should be able to sign certain
  messages.
\item[$\finsp: (\upk,\scred,\sig) \rightarrow \bin^*$.] Defines what utility
  value, derived from the user's public key, credentials, and produced signature
  (or a subset thereof), should be extractable by the group inspector.
\end{description}

@@

    \caption{Detailed oracles available in our model (1/2). Oracles for
      generating key material for users, issuers, and inspectors.}
    \label{fig:oracles1}
  \end{figure*}
}

@@

        \procedure{$\CHALb(\msg,
          (\uid_{0,1},\uattrs_{0,1},\dattrs_{0,1},\ldblbrace (\sgid,\cscid_{0,1},\tuattrs_{0,1},\tdattrs_{0,1}) \rdblbrace))$}{%
          \pcif \cuid_0 \notin \HU \lor \cuid_1 \notin \HU: \pcreturn \bot \\
          \pcif \feval(\msg,\CRED[\scid_0]) \neq \feval(\msg,\CRED[\scid_1]):
          \pcreturn \bot \\
          \csig \gets \Sign(\PRVUK[\uid_b],\msg,\uattrs_b,\dattrs_b, \\
          \hspace*{42pt}
          \ldblbrace (\GK[\sgid],\CRED[\scid_b],\tuattrs_b,\tdattrs_b) \rdblbrace) \\
          \CSIG \gets \CSIG \cup
          \lbrace (\csig,\msg,\cscid_b,\uattrs_b,\dattrs_b,\sgid) \rbrace \\
          \pcreturn \csig
        }

@@

\paragraph{Anonymity.} %
In group signatures, anonymity captures that no adversary must be able to learn,
from any group signature, the identity (e.g., member index) of its signer. In 
anonymous credentials, it requires that no adversary should learn anything about
the holder of a successfully shown credential, beyond that he owns a credential
attesting for the claimed function of the attributes it contains. In both GS and
AC, it is also typically required that
multiple signatures/presentations by the users are unlinkable. The approach to
formally state this property is in both cases frequently the same: the adversary
picks two (honest) users (or credentials in the AC case), the game randomly
chooses one of them, and lets the adversary request challenge
signatures/presentations from it. The adversary wins if it succeeds in guessing
which was the chosen user/credential better than guessing at random. In group
signatures, the game must also restrict the adversary from opening challenge
signatures. In anonymous credentials, the game must further constraint the
adversary to output credentials that have some common function of (a subset of)
its attributes.

In our notion of anonymity for \UAS, we need to merge the previous constraints.
Furthermore, a key difference with group signatures is that the game requires
the adversary to output credential identifiers, rather than user identifiers.
Specifically, this means that the adversary may actually output two credentials
that belong to the same user. Therefore, in some sense, the anonymity we get is
more general than that of group signatures. Moreover, in order to prevent
trivial wins by the adversary, we have to restrict that the challenge signatures
belong to a group that has been ``programmed'' with the same evaluation and
inspection functions, which is again a generalization over group signatures,
as discussed in the sequel. The formal specification of the anonymity game is
given in \figref{fig:exp-uas-anonb}.

@@

\begin{definition}{(Anonymity of \UAS)}
  We define the advantage \AdvAnon of $\adv$ against \ExpAnonb as
  $\AdvAnon=|\Pr\lbrack\ExpAnono(1^\secpar)=1\rbrack-
  \Pr\lbrack\ExpAnonz(1^\secpar)=1\rbrack|$.
  %
  An \UAS scheme satisfies anonymity if, for any p.p.t. adversary $\adv$,
  \AdvAnon is a negligible function of $1^\secpar$.
\end{definition}

\paragraph{Discussion on the generality of anonymity in \UAS schemes.} %
This notion of anonymity is more general than that of group signatures in the
sense that calls to the \INSPECT oracle reveal an arbitrary function of the
identity of the user -- which can certainly be the member index itself, as it
is frequent in group signatures, ore any other function computable from the
user public key, its credentials, and produced signature. \todo{Delve more in
  this.}

\paragraph{Traceability.} %
Traceability is one of the unforgeability-related properties in group
signatures. It captures that any signature accepted by \Verify needs to open
to one of the users that joined the group. While there is no traceability notion
in anonymous credentials, it is natural to map it to their unforgeability
property; if only because both require the issuer to be honest. Unforgeability
in anonymous credentials typically ensures that no adversary can get a verifier
to accept a credential presentation requiring a set of attributes that is not
contained in one of the credentials controlled by the adversary.

Our notion of traceability for \UAS is inspired by both. First, we restrict to
signatures that are accepted by \Verify. For any such signature, \Inspect has to
create a valid output (i.e., must not abort). The output of \Inspect has to be
accepted by \Judge too. Moreover, there must exist one user (either honest or
corrupt) who used his public key to obtain one or more credentials that allowed
him to produce a signature which, once inspected, returns \y.

\begin{figure}[htp!]
  \procedure{$\ExpTrace(1^\secpar)$}{%
    \parm \gets \Setup(1^\secpar) \\
    \pccomment{\todo{Right now, in \COGEN we may not know \osk! Then, the checks
        in the last part of the game may not be executable!}} \\
     (\sig,\dattrs,\msg) \gets
     \adv^{\lbrace \HI,\CO,\HU,\CU \rbrace \GEN,\OBTISS,\ISSUE,\SIGN,\INSPECT}
     (\parm) \\
     \pcreturn 1~\pcif \Verify(\sig,\msg,\dattrs,\GK[\sgid]) = 1~\land~( \\
     \pcind \pccomment{All \Obtain transcripts fail to open \sig} \\
     \pcind \nexists \gid \in \sgid:
     \Inspect(\PRVOK[\gid],\trans,\sig,\msg,\dattrs,\sgid) = \bot~\lor \\
     \pcind \pccomment{Such transcript exists, but ...} \\
     \pcind \exists\gid \in \sgid~\st~
     \Inspect(\PRVOK[\gid],\trans,\sig,\msg,\dattrs) =
     (\y,\iproof)~\land (\\
     \pcind \pcind \pccomment{... the proof is rejected by \Judge ...} \\
     \pcind \pcind \Judge(\y,\iproof,\sig,\msg,\dattrs,\GK[\sgid]) = 0~\lor \\
     \pcind \pcind \pccomment{... or \Judge accepts the proof, but inspect
       returns an invalid value} \\
     \pcind \pcind \nexists \uid \in \CU \cup \HU~\st~
     \y = \finsp(\PUBUK[\uid],\CRED[\uid],\sig) \\
     \pcind ) \\
     ) \\
     \pcreturn 0
  }
  \caption{Traceability experiment for \UAS schemes.}
  \label{fig:exp-uas-trace}
\end{figure}

\begin{definition}{(Traceability of \UAS)}
  We define the advantage \AdvTrace of $\adv$ against \ExpTrace as
  $\AdvTrace=\Pr\lbrack\ExpTrace(1^\secpar)=1\rbrack$.
  %
  A \UAS scheme satisfies traceability if, for any p.p.t. adversary $\adv$,
  \AdvTrace is a negligible function of $1^\secpar$.
\end{definition}

\paragraph{Discussion on the generality of traceability in \UAS schemes.} %
Note that this notion of traceability is, again, strictly more general than
both group signatures and anonymous credentials. On the group signatures side,
it all depends on the \Inspect function. If we set it to the traditional
inspection function, where it returns the identity (index) of the signer, then
our notion of traceability becomes exactly that of conventional group
signatures. On the other hand, \Inspect may be set to a non-injective function,
meaning that it can be possible that different user public keys, credentials and
signature tuples, may produce the same output. In that case, though, our notion
of traceability ensures that there must exist \emph{at least one} user capable
of issuing that signature.
%
On the anonymous credentials side, note that our definition allows that users
combine multiple credentials in a ``presentation''. However, in anonymous
credentials, it is frequent to require that all the attributes proven in a
presentation must be contained in a single credential. Note though that our
definition requires all credentials to be tied to the same user identity.
Therefore, it is a natural extension as (in the real life), it is very frequent
to combine multiple identifying documents to prove that you are entitled to
something. Moreover, any specific instantiation that restricts signatures
to be associated to just one credential directly leads to the previously
mentioned ``one credential per presentation'' rule.

@@

\begin{figure}[htp!]
  \procedure{$\ExpNonframe(1^\secpar)$}{%
    \parm \gets \Setup(1^\secpar) \\
     (\sig,\dattrs,\msg,\upk,\cred,\oproof) \gets
     \adv^{\lbrace\CI,\CO,\HU,\CU\rbrace\GEN,\OBTAIN,\SIGN}(\parm) \\
     \pcreturn 1~\pcif \Verify(\gpk,\sig,\dattrs,\msg) = 1 \land
     \OWNR[\cred] \in \HU \\%\land \cred \notin \CCRED~\land\\
     \pcind \Judge(\gpk,\upk,\cred,\oproof,\sig,\dattrs,\msg) = 1~\land \\
     \pcind \pccomment{\sig was not produced by \SIGN, or \upk and \cred do not match} \\
     \pcind (\sig \notin \SIG[\OWNR[\cred]] \lor \PUBUK[\OWNR[\cred]] \neq \upk) \\
     \pcreturn 0
  }
  \caption{Non-frameability experiment for \GSAC schemes.}
  \label{fig:exp-gsac-frame}
\end{figure}
